---
title: "Fetch & Process Results"
description: "Using the V1 API to download, reverse, and aggregate survey data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fetch & Process Results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval = FALSE}
library(formr)
```

The new `formr` API (V1) provides a streamlined pipeline for importing data. The API strictly separates **fetching** (getting data from the server) from **wrangling** (processing types, reversals, and scales), but the package provides a powerful convenience wrapper, `formr_api_results()`, to handle the entire process in one step.

## 1. Authentication

Before fetching data, ensure you are authenticated. If you have stored your keys (see [Getting Started](getting-started.html)), this is automatic.

```{r, eval = FALSE}
# Authenticate using stored credentials
formr_api_authenticate(host = "https://api.rforms.org", account = "dashboard") # or your custom URL and account name!
```

## 2. The One-Stop Solution (`formr_api_results`)

The primary function is `formr_api_results()`. By default, it performs a "Smart Fetch" that handles everything from download to psychometric scoring.

```{r, eval = FALSE}
# Fetch, reverse, aggregate, and join all data
df <- formr_api_results(run_name = "daily_diary")
```

### What does `formr_api_results` do by default?

1.  **Downloads** results and metadata for all surveys in the run.
2.  **Recognises** data types (converts "2023-01-01" to Dates, JSON arrays to lists, etc.).
3.  **Reverses** items ending in `R` (e.g., `neuro_3R`) based on item choices.
4.  **Aggregates** scales (calculates means for items with a shared stem).
5.  **Joins** all surveys into a single "wide" format (one row per session).

### Customizing the Fetch

You can toggle these behaviors using arguments if you want raw data or need to handle processing manually:

```{r, eval = FALSE}
# Get processed data, but strictly separated by survey (no join)
list_of_dfs <- formr_api_results("daily_diary", join = FALSE)

# Get raw data (types recognized, but NO reversing or scoring)
raw_data <- formr_api_results("daily_diary", compute_scales = FALSE)
```

------------------------------------------------------------------------

## 3. Advanced: The Manual Pipeline

If you used `compute_scales = FALSE` above, or if you are debugging why a specific scale isn't calculating correctly, you can use the lower-level processing functions manually. These rely on the survey's metadata (the item table).

### A. Get Metadata

Processing requires knowing which items are numeric, which are reversed (e.g., ends in `R`), and what the scale options are.

```{r, eval = FALSE}
# Fetch the survey structure (contains all item metadata)
metadata <- formr_api_survey_structure("daily_diary")
```

### B. Reverse Coding (`formr_api_reverse`)

This function automatically reverses items that end in `R` (e.g., `bfi_neuro_3R`). It uses the item's defined choices (e.g., 1 to 5) to calculate the reversal correctly: $New = (Max + Min) - Old$.

```{r, eval = FALSE}
# Apply reverse coding to raw results
df_reversed <- formr_api_reverse(results = raw_data, item_list = metadata)
```

*Result:* The column `bfi_neuro_3R` now contains reversed values. An attribute `reversed = TRUE` is added to the column for tracking.

### C. Aggregation (`formr_api_aggregate`)

This function calculates mean scores for items sharing a common "stem". For example, if you have `extra_1`, `extra_2`, and `extra_3R`, the function detects the stem `extra`, calculates the row-mean, and creates a new column `extra`.

```{r, eval = FALSE}
# Calculate scales (means)
df_scored <- formr_api_aggregate(results = df_reversed, item_list = metadata)
```

*Result:* A new column `extra` is added to the dataframe containing the mean of the items.

------------------------------------------------------------------------

## 4. Full Workflow Example

Here is a complete, reproducible snippet for a typical analysis script. Note how `formr_api_results` simplifies the process significantly compared to older versions of the package.

```{r, eval = FALSE}
library(formr)
library(dplyr)

# 1. Connect
formr_api_authenticate(host = "https://api.rforms.org") # or your custom URL!

# 2. Get Processed Data (Reversed & Scored)
run_name <- "daily_diary"
data <- formr_api_results(run_name)

# 3. Analyze
summary(data$bfi_neuro) # This scale was computed automatically!
# ...
```

## 5. Troubleshooting

### "My scale isn't calculating!"

Ensure your variable names follow the `stem_number` or `stem_numberR` convention (e.g., `neuro_1`, `neuro_2R`). `formr_api_aggregate` (and thus `formr_api_results`) relies on this to group items.

