---
title: "Getting Started"
description: "Installation, Authentication, and Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `formr` R package is the companion to the [formr.org](https://formr.org) survey framework. It allows you to:

1.  **Manage Projects:** Sync surveys and files between your computer and the server.
2.  **Fetch Data:** Download and process results for analysis.
3.  **Control Logic:** Use the API *within* a Run to check user progress or previous answers dynamically.

## Installation

You can install the package from GitHub:

```{r, setup, eval = FALSE}
if (!requireNamespace("remotes")) install.packages("remotes")
remotes::install_github("rubenarslan/formr")
```

The package is automatically installed within your formr server. Just use:

```{r, setup2, eval = FALSE}
library(formr)
```

## The "Two Ways of Connecting"

The package currently contains two ways of interacting with the formr-server.
It is helpful to know which one to use:

| Feature | **New API (V1)** | **Classic** |
| :--- | :--- | :--- |
| **Prefix** | `formr_api_*` | `formr_*` |
| **Authentication** | OAuth / Access Tokens | Email & Password |
| **Primary Use** | Project Management, Live Logic within your Run, Modern Data Fetching | Legacy Data Fetching |

**Recommendation:** Use the **API (`formr_api_*`)** for all new projects.

---

## Authentication

To use the API features (fetching results, managing files), you must first authenticate. 

### 1. Local Setup (One-time)

You can store your credentials securely in your system's keyring so you don't have to type them every time.

#### 1.1 Save your API Credentials

1.  Log in to formr.org (or your instance).
2.  Go to **Account > API Credentials** (the "API" tab on the account page). You can hold
    several credentials side by side — each one with its own scopes and run allowlist.
    A common pattern is one narrow read-only credential for a dashboard, plus a separate
    broader credential for a cron job.
3.  Give the credential a **label** (e.g. `dashboard`, `cron-2026`). The label is only
    used on the credentials page to tell credentials apart; it must be unique within
    your account. You will pass the same string to `formr_store_keys(..., account = "...")`
    in step 6 so the R package can keep multiple credentials per host straight in
    your keyring.
4.  Pick which **scopes** the credential should grant. Each scope is a verb on a resource:

    | Scope | What it grants |
    | :--- | :--- |
    | `user:read` / `user:write` | Read / update your account profile |
    | `survey:read` / `survey:write` | Read survey definitions / upload + edit them |
    | `run:read` / `run:write` | Read run metadata / create + update + delete runs |
    | `session:read` / `session:write` | Read participant sessions / create + advance them |
    | `data:read` | Read participant response data |
    | `file:read` / `file:write` | Download / upload files attached to runs |

    A token issued under a credential can only call endpoints whose required scopes it carries. A credential with only `run:read` will succeed on `GET /v1/runs/{name}` and return 403 on `PATCH /v1/runs/{name}`.

5.  Optionally **restrict the credential to specific runs**. Leave the run picker empty to allow this credential to act on all of your runs; tick one or more runs to limit it. A run-restricted credential also implicitly restricts which surveys it can touch — only surveys that are part of one of the allowlisted runs are reachable.

6.  Click **Create credential**. The Client ID and Client Secret are shown **once**. Copy both immediately — the server stores only a SHA-256 hash, so a forgotten secret has to be rotated, not recovered.

7.  Run the following code **once** in your R console:

```{r, store_keys, eval = FALSE}
# Store your credentials once
# This saves them securely in your OS credential store
formr_store_keys(
  host = "https://api.formr.org",
  client_id = "YOUR_CLIENT_ID",
  client_secret = "YOUR_CLIENT_SECRET",
  # Optional but recommended when you have more than one credential:
  # pass the same string you used as the credential's label on the
  # server. This lets you switch between credentials by name when you
  # call formr_api_authenticate(account = "...").
  account = "dashboard"
)
```

#### 1.2 Save your regular User Credentials (Classic)

If you plan to use the "Classic" functions (e.g. `formr_results()`, `formr_raw_results()`), you need to store your email and password. This allows you to create a shorthand name for your account to log in securely without putting passwords in your script.

1.  Run the following code **once** in your R console.
2.  You will be prompted to enter your email and password (and optionally a 2FA secret) securely.

```{r, store_keys_classic, eval = FALSE}
# Store your email/password under a shorthand name (e.g. "main_account")
formr_store_keys("main_account")
```

### 2. Authenticating in Scripts

Once your keys are stored, you simply authenticate at the start of your session:

#### 1.1 Thru the API

```{r, auth_local, eval = FALSE}
# Automatically finds your stored keys
formr_api_authenticate(host = "https://api.formr.org", account = "dashboard") # or your custom API-URL + account name!

# After authentication, you can see which scopes the credential carries:
formr_api_session()$scope
#> [1] "run:read run:write survey:read"
```

If you call a function whose endpoint needs a scope you don't have, the package surfaces the server's 403 alongside a hint at the bottom pointing at `admin/account#api` — the same place to fix it.

#### 1.2 With your User Credentials (Classic)

Once stored, you can connect in your scripts using this shorthand:

```{r, connect_classic, eval = FALSE}
# Connect using the stored credentials
formr_connect("main_account")
```

### 3. Authenticating *Inside* formr Runs

You can also use the API functions within the R-Code of your actual survey run (e.g., to look up previous results). In this environment, you do not need credentials; the system provides a temporary context. Just run:

```{r, auth_run, eval = FALSE}
# Inside a formr Run, simply call:
formr_api_authenticate() # The package detects that it's running inside a Run and uses the temporary context provided by the server.
```

---

## Workflow Examples

### Project Management (Push & Pull)

The new API allows you to work on your survey structure and files locally and sync them to the server.
Learn more about how this can save you time in the [Manage your Projects](manage-your-projects.html) Tutorial.

```{r, push_pull, eval = FALSE}
# Download a project (surveys and files) to your local folder
formr_api_pull_project("daily_diary")

# Upload changes back to the server
formr_api_push_project("daily_diary")
```

### Fetching Results

The `formr_api_results()` function is the modern way to get data. It automatically recognizes data types, reverses items, and aggregates scales. Learn how it works in the [Fetch & Process Results](fetch-and-process-results.html) Tutorial.

```{r, results, eval = FALSE}
# Fetch and process
df <- formr_api_results("daily_diary")
```

---

## Token Management & Security

When you run `formr_api_authenticate()`, it provides you with an Access Token.

* **Validity:** The token is valid for **one hour**. You will need to rerun authentication after this period.
* **On the Server:** When running code *inside* a formr survey, the token is invalidated automatically when the request finishes.
* **Local Device:** When running code on your laptop, the token remains valid until it expires or you revoke it.

At the end of your analysis session, you may revoke the token manually:

```{r, logout, eval = FALSE}
formr_api_logout()
```