---
title: "Running R Inside Your formr Study"
description: "Use the R package from calculate items, showif conditions, and inline labels within a run"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running R Inside Your formr Study}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

If you haven't already, read the [Getting Started](getting-started.html) guide for authentication basics.

## 1. Why Run R Inside Your Study?

Running R code inside your formr study has always been possible — calculate
items, showif conditions, and inline labels can already draw on the current
participant's own data. The V1 API **broadens** this to data from *other*
participants. A calculate item can now ask "what did everyone else answer?"
and branch, display, or store based on the answer. Here is what becomes
possible:

- **Adapt in real time.** Check what other participants have answered so far and
  branch, skip, or compute norms on the fly.
- **Balance experimental groups automatically.** Count completions per condition
  and route each new participant to the smaller cell, compensating for differential
  attrition without manual intervention.
- **Synchronise participants.** Poll the database for incoming sessions and
  advance a whole cohort at once — enabling live dyadic tasks, focus groups,
  or team exercises inside an asynchronous survey framework.

All of these patterns are built on the same small set of tools: authentication,
context variables, and a single data-fetching function. The walkthroughs in
§4–§7 show each pattern from start to finish.

These are just starting points — since you can run arbitrary R code, any
server-side logic that benefits from cross-session data or further API functions is fair game.

## 2. Where the API Code Goes

The V1 API is called from **calculate items** — hidden fields that run R on
the server between surveys. The results are then displayed or acted upon by
two other mechanisms that already existed in formr.

### A. Calculate Items (API Entry Point)

A calculate item evaluates an R expression when the participant reaches it.
The last value is stored in the session data and can be used by later units
(showif conditions, labels, other calculate items). This is where
`formr_api_authenticate()` and `formr_api_fetch_results()` go.

```{r, eval = FALSE}
# Inside a calculate item's "value" field:
formr_api_authenticate()

past <- formr_api_fetch_results(.formr$run_name, item_names = "score", join = TRUE)
mean(past$score, na.rm = TRUE)
```

Use calculate items for any logic that needs cross-session data: computing
norms, generating tokens, processing JSON, or counting completions per
experimental cell.

### B. Inline R in Labels (Display + Fetch)

Labels render text and plots to the participant. They can call API functions
just like calculate items — `formr_api_authenticate()` and
`formr_api_fetch_results()` work here too.

````{r, eval = FALSE}
# In a note item's label — fetch and display in one step:
# ```{r, echo=FALSE, results='asis'}
# formr_api_authenticate()
# scores <- formr_api_fetch_results(.formr$run_name,
#   item_names = "engagement", join = TRUE)
# cat("The sample mean is ", mean(scores$engagement, na.rm = TRUE), ".")
# ```
````

Use `echo=FALSE` to hide the code and `results='asis'` to print raw output.
You can also render `ggplot2` charts this way (see the Group Norms walkthrough
in §5).

As a style choice, you can separate fetch logic into a calculate item and keep
labels minimal — this makes studies easier to debug. But nothing prevents you
from doing both in one label, as the Group Norms walkthrough in §5 does.

### C. Inline R in Choices (Display)

Choice labels can also display dynamic content, useful when a dropdown menu
should reflect live database contents.

````{r, eval = FALSE}
# Inside a mc_multiple choice option's label — fetch and display in one step:
# ```{r}
# formr_api_authenticate()
# posts <- formr_api_fetch_results(.formr$run_name, item_names = "title", join = TRUE)
# posts$title[1]
# ```
````


## 3. Your Toolkit

Every piece of server-side R code needs the same few ingredients.

### Authentication

Inside a run, credentials are injected automatically. Just call:

```{r, eval = FALSE}
formr_api_authenticate()
```

The package detects `.formr$access_token` and `.formr$host` set by the server.
The token is valid for the duration of the request and is revoked when the
request finishes.

### Run Context

Two hidden variables are always available:

| Variable | What it holds |
|---|---|
| `.formr$run_name` | The name of the current run (e.g. `"daily_diary"`) |
| `survey_run_sessions$session` | The current participant's session code |

```{r, eval = FALSE}
run_name <- .formr$run_name
user_session <- survey_run_sessions$session
```

Use these to fetch the right data and associate new data with the right
session, making your code portable across runs.

### Fetching Data from Other Surveys

The function for reading data *from within a run* is `formr_api_fetch_results()`.
It differs from `formr_api_results()` in important ways:

| | `formr_api_results()` | `formr_api_fetch_results()` |
|---|---|---|
| Auto-reverses items | Yes | No |
| Auto-computes scales | Yes | No |
| Returns processed data | Yes | No |
| `item_names` filter | No | Yes |
| Default `run_name` | `.formr$run_name` | `.formr$run_name` |
| Default `join` | `TRUE` | `FALSE` |

Inside a run, you almost always want `formr_api_fetch_results()` — raw data
without transformations is safer when you process it yourself.

```{r, eval = FALSE}
data <- formr_api_fetch_results(.formr$run_name,
  item_names = c("name", "age", "score"), join = TRUE)
```

Always specify `item_names` to keep requests fast. The result is a tibble with
one row per session and a column per requested item (plus a `session` column).
If the same item name appears in multiple surveys, the survey name is prefixed.

### The `current()` Shorthand

In showif conditions and value expressions, formr repeats items within a
session. The helper `current(x)` returns the **most recent submission** of
an item — the last element of the vector, which is always the current
session's value:

```{r, eval = FALSE}
# In a showif condition — check the current selection
current(menu_survey$choice) == "option_a"

# In a calculate item — capture the latest input
current(my_survey$text_input)
```

This is cleaner than the equivalent base-R `x[length(x)]` pattern and makes
your intent explicit. See `?current` for details.

---

## 4. Walkthrough: Participant Counter

The simplest complete example: greet each participant by their number in the
study, using cross-session data.

### Run Structure

| Position | Type | Name | What it does |
|---|---|---|---|
| 10 | Survey | `register` | Collects participant's name |
| 20 | Calculate | `participant_count` | Counts all registrations so far |
| 30 | Survey | `welcome` | Shows "You are participant #N" |

### Calculate: `participant_count`

```{r, eval = FALSE}
formr_api_authenticate()

past <- formr_api_fetch_results(.formr$run_name, item_names = "name", join = TRUE)

if (nrow(past) > 0) nrow(past) + 1 else 1
```

### Survey: `welcome`

A note item with this label:

````{r, eval = FALSE}
# ```{r, echo=FALSE, results='asis'}
# cat("## Welcome, Participant #", participant_count, "\n\n", sep = "")
# cat("Please proceed with the study.")
# ```
````

## 5. Walkthrough: Real-Time Group Norms

**Problem:** A single score tells a participant nothing. Showing how they
compare to the current sample (descriptive norms) increases engagement and
provides real value.

**Research context example:** Occupational burnout surveys where participants see
their score plotted against the organisational distribution in real time.

### Run Structure

| Position | Type | Name | What it does |
|---|---|---|---|
| 10 | Survey | `burnout` | Contains a regular item called `engagement` |
| 20 | Survey | `feedback` | Note item whose label fetches, plots, and displays |

The `burnout` survey has a regular item called `engagement` where the
participant enters their score. The feedback label does everything in one
step — no separate calculate item needed, no data stored between units.

### Survey: `feedback` (label)

A note item whose label fetches all engagement scores via the API and renders
a ggplot comparing the current participant against the sample distribution:

````{r, eval = FALSE}
# ```{r, echo=FALSE, results='asis', fig.width=6, fig.height=3}
# library(ggplot2)
# formr_api_authenticate()
#
# # Fetch all participants' engagement scores
# all_scores <- formr_api_fetch_results(.formr$run_name,
#   item_names = "engagement", join = TRUE)
#
# # Current participant's own score — local, no API needed
# my_engagement <- current(burnout$engagement)
#
# ggplot(all_scores, aes(x = engagement)) +
#   geom_density(fill = "grey70") +
#   geom_vline(xintercept = my_engagement, colour = "red", linewidth = 1) +
#   labs(
#     title = "Your engagement score vs. the organisation",
#     subtitle = paste0("Sample: ", nrow(all_scores), " colleagues"),
#     x = "Engagement", y = ""
#   ) +
#   theme_minimal()
# ```
````

Note two patterns worth reusing:

1. The current participant's own score (`current(burnout$engagement)`) is
   available locally from the session — no API roundtrip needed.
2. The API call (`formr_api_fetch_results`) only fetches what the label
   cannot already see: *other* participants' data. This is one roundtrip,
   one OpenCPU session, and nothing stored in the database beyond what
   the `burnout` survey already saves.

---

## 6. Walkthrough: Dynamic Group Balancing

**Problem:** In field experiments, attrition often differs between conditions.
Static random assignment at the start produces unequal cell sizes by the end.
Manually monitoring and rebalancing is tedious and error-prone.

**Solution:** Count completed sessions per condition on every new entry and
route the participant to the currently smaller group.

### Run Structure

| Position | Type | Name | What it does |
|---|---|---|---|
| 10 | Survey | `intake` | Baseline demographics |
| 20 | Calculate | `pick_condition` | Fetches prior completions, picks smaller group |
| 30 | Survey | `intervention_a` | Treatment module A (shown if condition == "A") |
| 40 | Survey | `intervention_b` | Treatment module B (shown if condition == "B") |

### Calculate: `pick_condition`

```{r, eval = FALSE}
formr_api_authenticate()

# Fetch the condition assignments from all completed sessions
past <- formr_api_fetch_results(.formr$run_name,
  item_names = "assigned_condition", join = TRUE)

count_a <- sum(past$assigned_condition == "A", na.rm = TRUE)
count_b <- sum(past$assigned_condition == "B", na.rm = TRUE)

# Assign to the smaller group; break ties randomly
if (count_a <= count_b) "A" else "B"
```

### Showif conditions

Position 30 (`intervention_a`) showif:
```
current(pick_condition) == "A"
```

Position 40 (`intervention_b`) showif:
```
current(pick_condition) == "B"
```

---

## 7. Walkthrough: Synchronising with a Waiting Room

**Problem:** Live dyadic tasks, focus groups, and team exercises require
multiple participants to start a module simultaneously. Asynchronous survey
frameworks let everyone progress at their own pace.

**Solution:** Trap early arrivals in a refresh loop, then advance the whole
cohort at once via the API.

### Run Structure

| Position | Type | Name | What it does |
|---|---|---|---|
| 10 | Survey | `lobby` | Intake survey |
| 20 | Survey | `waiting_room` | Auto-refreshing hold page (submit: 2000 ms) |
| 30 | SkipBackward | `loop_back` | Returns to position 20 |
| 40 | Survey | `dyadic_task` | The live interaction — only visible after release |

The waiting room survey has a single hidden submit button configured with
`submit: 2000` in its survey settings, causing it to re-submit every 2
seconds. A SkipBackward unit (position 30) immediately sends the session
back to position 20, creating a continuous loop.

An **administrator trigger** (run manually or on a cron schedule) polls for
queued sessions and releases them:

```{r, eval = FALSE}
# Admin trigger — run outside the study, on your local machine or a cron job
formr_api_authenticate(host = "https://api.rforms.org", account = "admin")

# Find sessions currently at the waiting room (position 20)
queued <- formr_api_sessions("my-run-name", active = TRUE)
waiting <- queued$session[queued$position == 20]

if (length(waiting) >= 2) {
  formr_api_session_action("my-run-name",
    session_codes = waiting, action = "move_to_position", position = 40)
  message("Released ", length(waiting), " participants to the dyadic task.")
}
```

After release, sessions land directly on position 40 (the `dyadic_task`
survey), bypassing the SkipBackward loop.

### Participant experience

1. Complete the lobby survey.
2. Land on the waiting room — the page auto-re-submits every 2 s,
   keeping them in the loop.
3. When the admin trigger finds enough participants, it moves them past
   the SkipBackward to the live task.

---

## 8. Patterns for Robust Code

The walkthroughs above keep code minimal for clarity. When you adapt them
to a real study, a few defensive habits will save you debugging time — code
running inside formr runs on the server and a single unhandled error can
break a participant's flow.

**Handling JSON data.** formr handles JSON in three ways:

1. **Automatic:** When a calculate item returns a named list (like
   `list(my_score = ..., sample_n = ...)`), formr serialises it to JSON
   and stores it. You never write `toJSON()` for this — it just works.
2. **Persistence:** For complex state that must survive across OpenCPU
   requests (e.g., a balancing matrix or network adjacency list), use
   `jsonlite::toJSON(x, auto_unbox = TRUE)` to store it explicitly as a
   session variable, and `jsonlite::fromJSON()` to read it back in a
   later calculate item.
3. **Parsing responses:** API columns sometimes contain nested JSON
   (e.g., multi-select choices). Always wrap parsing in a `tryCatch`:

```{r, eval = FALSE}
safe_parse_json <- function(x) {
  tryCatch(
    jsonlite::fromJSON(x),
    error = function(e) list()
  )
}
```

**Guard against empty results.** Before processing fetched data, check
that it has rows:

```{r, eval = FALSE}
data <- formr_api_fetch_results(.formr$run_name, item_names = "score", join = TRUE)

if (nrow(data) == 0 || all(is.na(data$score))) {
  result <- 0
} else {
  result <- max(data$score, na.rm = TRUE)
}
```

**Use `item_names`.** Always specify the exact columns you need when
calling `formr_api_fetch_results()` inside a run. This keeps requests
fast and avoids pulling unnecessary data.

---

## Next Steps

- The [Fetch & Process Results](fetch-and-process-results.html) vignette
  covers the downstream analysis pipeline.
- The `?formr_api_fetch_results` and `?formr_api_authenticate` help pages
  have the full parameter details.
- See `?current`, `?first`, and `?last` for the other shorthand helpers
  available inside a run.
