```{r eval=TRUE,echo=FALSE}
if (!exists("indent")) {
	indent = '#' # ugly hack so _regression_summary can be "spun" (variables included via `r ` have to be available)
}
if (exists("testing")) {
	results = data.frame()
}

users = dplyr::n_distinct(results$session)
finished_users = dplyr::n_distinct(results[!is.na(results$ended),]$session)
rows_per_user = nrow(results)/users
repeated_survey = rows_per_user > 1
duration = results %>% filter(!is.na(ended)) %>% 
	dplyr::mutate(duration = as.double(ended - created, unit = "mins"))

started = sum(!is.na(results$modified))
only_viewed = sum(is.na(results$ended) & is.na(results$modified))
```

`r indent`# Survey overview

`r finished(results)` completed rows, `r started` who entered any information, `r only_viewed` only viewed the first page. There are `r expired(results)` expired rows (people who did not finish filling out in the requested time frame). In total, there are `r nrow(results)` rows including unfinished and expired rows. 

There were `r users` unique participants, of which `r finished_users` finished filling out at least one survey.

`r round(rows_per_user,2)` rows per user code. `r ifelse(repeated_survey, "A repeated survey")`. `r ifelse(repeated_survey & round(rows_per_user) == 1, "Oddly there seem to be almost no users with several rows. Maybe something went wrong and a few rows are duplicated by accident?", "")`


```{r}
if (repeated_survey) {
	overview = results %>% dplyr::group_by(session) %>% 
		dplyr::summarise(
			n = sum(!is.na(session)),
			expired = sum(!is.na(expired)),
			ended = sum(!is.na(ended))
		) %>% 
		tidyr::gather(key, value, -session)
	if (length(unique(dplyr::filter(overview, key == "expired")$value)) == 1) {
		overview = dplyr::filter(overview, key != "expired")
	}
	print(
		ggplot2::ggplot(overview, ggplot2::aes(value, ..count..)) + ggplot2::geom_bar() + ggplot2::facet_wrap(~ key, nrow = 1)
	)
}
```

The first session started on `r min(results$created)`, the last session on `r max(results$created)`. People took on average `r mean(duration$duration)` minutes (median `r median(duration$duration)`) to answer the survey.

```{r}
high_vals = sum(median(duration$duration) + 4*mad(duration$duration) < duration$duration)
ggplot2::qplot(duration$duration, binwidth = 0.5) + ggplot2::scale_x_continuous(paste("Duration (in minutes), excluding", high_vals, "values above median + 4*MAD"), limits = c(NA, median(duration$duration) + 4*mad(duration$duration)))
```



`r indent`# Scales

```{r results='asis'}
vars = names(results)
for (i in seq_along(vars)) {
	var = vars[i]
	cat(var,"\n", file = "~/Downloads/var.log", append = T)
	scale = results[[ var ]]
	scale_info = attributes(scale)
		if ( !is.null(scale_info)) {
		if (exists("scale", scale_info)) {
		cat(codebook_component_scale( scale ))
		}
	}
}
```

`r indent`# Single items

```{r results='asis'}
for (i in seq_along(vars)) {
	var = vars[i]
	cat(var,"\n", file = "~/Downloads/var.log", append = TRUE)
	item = results[[ var ]]
	item_info = attributes(item)
	if ( !is.null(item_info)) {
		if (exists("part_of_scale", item_info) || exists("scale", item_info)) {
			next
		} else if (exists("item", item_info)) {
			cat(codebook_component_single_item( item ))
		} else {
			cat(codebook_component_fallback( item, var))
		}
	}
}
```


`r indent`# Missingness report {.tabset .tabset-sticky}

There were `r finished(results)` rows completed, `r started` who entered any information, `r only_viewed` only viewed the first page. There are `r expired(results)` expired rows (people who did not finish filling out in the requested time frame). 

### Missingness finishers

among those who finished the survey

```{r}
missingness = missingness_patterns(
	dplyr::select(dplyr::filter(results, !is.na(ended)), -expired, -ended) )
```
```{r}
pander::pander(missingness)
```



### Missingness starters

among those who started the survey

```{r results='markup'}
missingness = missingness_patterns(
	dplyr::filter(results, !is.na(modified)) )
```
```{r}
pander::pander(missingness)
```
